---
title: "❄📊️☃️️ Anchorage Snowfall"
format: 
  html:
    toc: true
    embed-resources: true
    theme: zephyr
date: "`r Sys.Date()`"
execute: 
  warning: false


---



<!-- {{include-in-header }} -->


<!-- <link rel="preconnect" href="https://fonts.googleapis.com"> -->
<!-- <link rel="preconnect" href="https://fonts.gstatic.com" crossorigin> -->
<!-- <link href="https://fonts.googleapis.com/css2?family=Fira+Code:wght@300..700&family=IBM+Plex+Mono:ital,wght@0,200;0,400;0,600;0,700;1,400&family=Inter:wght@600;800&display=swap" rel="stylesheet"> -->


```{css, echo = FALSE}
* {font-family: "Fira Code";

}

h1,h2,h3,h4,h5,h6{

font-family: "Inter"

}

```

## Anchorage Ted Stevens International Airport Snowfall 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = F, fig.asp = .8, fig.width = 10)




# 
# install.packages('readr')
# install.packages('dplyr')
# install.packages('readr')
# install.packages('stringr')
# install.packages('janitor')
# install.packages('jsonlite')
# install.packages('ggplot2')
# install.packages('lubridate')
# install.packages('directlabels')
# install.packages('purr')
# install.packages('gghighlight')
# install.packages('plotly')
# install.packages('knitr')
# install.packages('httr')
# install.packages("curl")



library(tidyr)
library(dplyr)
library(readr)
library(stringr)
library(janitor)
library(jsonlite)
library(ggplot2)
library(lubridate)
library(directlabels)
# library(dataRetrieval)
library(purrr)
library(gghighlight)
library(plotly)
# library(ggthemr)
# library(RColorBrewer)
library(knitr)
library(httr)


# 
# setwd("projects/iditarod-rest")
# setwd("/")
options(scipen=999)
getwd()


# source("../ggplot_theme/ben_theme.R")


```




<!-- ## Read in Historic Data -->

```{r}

hx_files <- list.files("data/historic",full.names = T )
hx_files <- as.list(hx_files)


hx <- lapply(hx_files, read_csv) %>% bind_rows()
hx <- hx %>% clean_names()

hx <- hx %>% select(-name, -mdpr, -dapr, -prcp)

hx <- hx %>% relocate(date)
hx <- hx %>% filter (date<'2024-01-01')

# 
# write_csv(hx, "hx.csv")
# write_json(hx, "hx.json")

```



```{r}


WXTOKEN <- Sys.getenv("WXTOKEN")

```


```{r}





##old
daily_url_2024 <- "https://www.ncei.noaa.gov/cdo-web/api/v2/data/?datasetid=GHCND&stationid=GHCND:USW00026451&startdate=2024-01-01&enddate=2024-12-31&limit=999&datatypeid=SNOW&datatypeid=SNWD&units=standard"


daily_raw_24 <- httr::GET(daily_url_2024,add_headers( "token" = WXTOKEN
) )


# print("daily_raw_24")
# print(daily_raw_24)

daily_content_24 <- content(daily_raw_24)


# print("dialy content 24")
# print(daily_content_24)


daily_df_24 <- bind_rows(daily_content_24$results)


# print("daily_df_24")
# print(daily_df_24)

snow_24 <- daily_df_24 %>% filter (datatype %in% c("SNOW", "SNWD"))
snow_24_wide <- snow_24 %>% pivot_wider( names_from   = datatype)
# snow_24_wide <- snow_24_wide %>% select(-resultset,-attributes)
snow_24_wide <- snow_24_wide %>% select(-attributes)
snow_24_wide <- snow_24_wide %>% clean_names()




daily_url_2025 <- "https://www.ncei.noaa.gov/cdo-web/api/v2/data/?datasetid=GHCND&stationid=GHCND:USW00026451&startdate=2025-01-01&enddate=2025-12-31&limit=999&datatypeid=SNOW&datatypeid=SNWD&units=standard"



daily_raw_25 <- httr::GET(daily_url_2025,add_headers( "token" = WXTOKEN
) )

daily_content_25 <- content(daily_raw_25)

daily_df_25 <- bind_rows(daily_content_25$results)

snow_25 <- daily_df_25 %>% filter (datatype %in% c("SNOW", "SNWD"))
snow_25_wide <- snow_25 %>% pivot_wider( names_from   = datatype)
# snow_25_wide <- snow_25_wide %>% select(-resultset,-attributes)
snow_25_wide <- snow_25_wide %>% select(-attributes)
snow_25_wide <- snow_25_wide %>% clean_names()



```



```{r}


snow <- rbind(hx, snow_24_wide, snow_25_wide)

snow <- snow %>% distinct()



snow <- snow %>% mutate(date = date(date))
snow <- snow %>% mutate(year = year(date))
snow <- snow %>% mutate(month= month(date))
snow <- snow %>% mutate(previous_year = year-1)
snow <- snow %>% mutate(next_year = year+1)




# add julian date
snow <- snow %>% mutate(date_julian = yday(date))

## date 2000
snow <- snow %>% mutate(date_string = paste0(date))
snow <- snow %>% mutate(date_part = substr(date, 6,10))
snow <- snow %>% mutate(date_month = substr(date, 6,7))

snow  <-  snow %>% filter(month %in% c(1,2,3,4,5,9,10,11,12))
snow <- snow %>% select(-station)

# snow <- snow %>% mutate(date_2000 = ifelse(date_month %in% c('10', '11', '12'), paste0("2000-", date_part), paste0("2001-", date_part) ))
# snow <- snow %>% mutate(season = ifelse(date_month %in% c('9','10', '11', '12'), "fall", "spring"))
snow <- snow %>% mutate(season = ifelse(month %in% c(9,10,11,12), "fall", "spring"))
# snow <- snow %>% mutate(date_2000 = paste0("2000-", date_part))


snow <- snow %>% mutate (winter_season = ifelse(season=="fall", paste0(year, "-", next_year ), paste0(previous_year,"-", year ) ))

##establish a start date for each year. 


snow <- snow %>% mutate(snow_year_start_year = ifelse(season =="fall", year, previous_year))

snow <- snow %>% mutate(snow_year_start_date = paste0(snow_year_start_year, "-09-01"))


##caluclate the number of DAYS since the statrt date

snow <- snow %>% mutate(snow_year_interval = interval(start=snow_year_start_date, end=date ))

snow <- snow %>% mutate(snow_year_period = as.period(snow_year_interval, units="days" ))

snow <- snow %>% mutate(snow_year_day_dt = difftime( date, snow_year_start_date, units="days") )



# snow <- snow %>% mutate(snow_year_days= days(snow_year_period) )



# snow <- snow %>% mutate(date_2000_date = ymd(date_2000))



### calculate running snowfall total: 



snow <- snow %>% arrange(date)


## add decades 

snow <- snow %>% mutate (decade=(year %/% 10) *10)

## find NAs

snow_na <- snow %>% filter (is.na(snow))
snow <- snow %>% filter (!is.na(snow))


snow <- snow %>% group_by(winter_season) %>% mutate (snow_rolling = cumsum(snow)) %>% ungroup()

# 2023-2024

```

```{r}
latest_date <- max(snow$date)

## top 10 winters


snow_summary <- snow %>% group_by(winter_season) %>%  summarize(max_cumulative = max(snow_rolling), min_cumulative = min(snow_rolling), decade = first(decade)) %>% ungroup()


snow_top_10 <- snow_summary %>% top_n(10, max_cumulative) %>% arrange(-max_cumulative)

snow_smallest_10 <- snow_summary %>% top_n(10, -max_cumulative) %>% arrange(-max_cumulative)

# snow_smallest_10 <- snow_summary %>% slice_min( order_by=min_cumulative, n=10) %>% arrange(min_cumulative)

```


<!-- The latest date in the snow data is `r paste0(latest_date)` -->

📅This report was run `r today()`. 

🔗The source of the snowfall data is the *[NOAA Nation Centers for Environmental Information](https://www.ncei.noaa.gov/cdo-web/)* Climate Data Online (CDO) page. The site is shown here is Ted Stevens Anchorage International Airport. 

## 2024-2025 in Anchorage cumulative snowfall compared to winters since 1952.

```{r}

ggplot(snow %>% filter(month!= 99))+
  geom_line(aes(x=snow_year_day_dt, y=snow_rolling, group=winter_season, color=winter_season), show.legend = F)+
  # gghighlight(winter_season == '2023-2024')+
  gghighlight(winter_season == '2024-2025')+
  ggtitle("Winter 2024-2025")+
  theme_minimal()

```



## 2024-2025 in Anchorage cumulative snowfall compared to winters since 1952 - Distribution

```{r}

ggplot(snow_summary)+
  geom_histogram(aes(x=max_cumulative, show.legend = F))+
  # gghighlight(winter_season == '2023-2024')+
  gghighlight(winter_season == '2024-2025')+
  theme_minimal()

  # ggtitle("Winter 2024-2025")+

```


```{r}

ggplot(snow_summary)+
  geom_histogram(aes(x=max_cumulative, show.legend = F))+
  # gghighlight(winter_season == '2023-2024')+
  gghighlight(winter_season == '2024-2025')+
  theme_minimal()+
  facet_wrap(~decade)

```



## Largest Top 10 Years of Snowfall


```{r}
kable(snow_top_10)

```
## Smallest Top 10 Years of Snowfall

```{r}
kable(snow_smallest_10)

```







## Anchorage cumulative snowfall compared to winters since 1952 - Interactive Plot
```{r}

all_years_plot <- ggplot(snow %>% filter(month!= 5))+
  geom_line(aes(x=snow_year_day_dt, y=snow_rolling, group=winter_season), show.legend = F, color="gray")+
  theme_minimal()+
  theme(legend.position='none') 


ggplotly(all_years_plot, tooltip = c("snow_year_day_dt", "snow_rolling", "winter_season"))

```


## Snow Day Patterns 2023-2024 and 2024-2025

```{r}



ggplot()+
  geom_col(data = snow %>% filter(winter_season %in% c("2024-2025", "2023-2024")), aes(x=snow_year_day_dt, y=snow, fill=winter_season))+ facet_wrap(~winter_season, ncol=1)+
  theme_minimal()

```



## 2024-2025 Anchorage snow depth compared to winters since 1952.

```{r}

ggplot(snow)+
  geom_line(aes(x=snow_year_day_dt, y=snwd, group=winter_season, color=winter_season), show.legend = F)+
    gghighlight(winter_season == '2024-2025')+
  theme_minimal()+
   scale_color_brewer(palette = "Dark2")

```



<!-- ## 2023-2024 Anchorage snow depth by decades since 1952. -->

```{r}

# ggplot(snow %>% filter(month!= 5))+
  # geom_line(aes(x=snow_year_day_dt, y=snow_rolling, group=winter_season, color=decade), show.legend = F)+
  # theme_minimal()
  # scale_color_brewer(palette = "PuBu")

```

## 2023-2024 Anchorage snow cumulative compared to the top 10 winters

```{r}

ggplot(snow %>% filter (winter_season %in% snow_top_10$winter_season))+
  geom_line(aes(x=snow_year_day_dt, y=snow_rolling, group=winter_season, color=winter_season), show.legend = F)+
    gghighlight(winter_season == '2023-2024')+
  theme_minimal()

```





## Cumulative Snowfall- Decade Facets 

```{r}

ggplot(snow %>% filter(month!= 5))+
  geom_line(aes(x=snow_year_day_dt, y=snow_rolling, group=winter_season, color=factor(decade)), show.legend = F)+
  facet_wrap(~decade)+
  theme_minimal()

```
